---
title: "Lab 3 Key"
author: "Nathaniel Grimes"
format: html
---
## Load in data and packages

```{r}
library(tidyverse)
library(tidymodels)
library(here)

t_df<-read_csv(here("data", "titanic_survival.csv"))
```
1. What's your research Q
2. Gather data
3. Develop a hypothesis
4. Define methodology
5. Look at data / EDA
6. Clean data / determine which variables to keep

## ML Framework

Write a description of how you go about applying machine learning problems. If you drew a diagram, share with a neighbor and discuss.

1. Define metrics for success
2. Data split / partitioning
3. Build Models
  a. Create model options by defining what variables you want to include
  b. Data preprocessing
  c. train model
  d. tune our model
  e. evaluate perf
  f. model selection

4. Finalize the model (aka train it on all the data)
5. Interpret results/visualize/communicate

.Get to C-fold validation

Using a specified portion of a known data set to train a model's predictive ability, then testing and assessing that model's predictive accuracy


## Tidymodels Overview

The creators of `tidyverse` have created a new package called `tidymodels` that is designed to make machine learning more accessible to R users. The package is designed to work with the `tidyverse` and `tidydata` principles. 


## Defining a Research Question

What are we trying to solve? The crucial step of any scientist that can take years to define and perfect.

**What factors led to the survival of passengers on the Titanic?**

How will go about solving our question? Use a classification algorithm to predict the survival of passengers on the Titanic. Interpret the variables that control observed outcomes.

In real life, we would have to go out and collect the data. Today, we will use the `titanic` dataset from the `titanic` package. 


## Data Exploration/Cleaning

Take 15 minutes to explore the data. Are there any immediate changes to the data that we need to change? What relationships can you see through graphs? What variables could be of interest to predict survival of passengers on the Titanic?

need to understand some of these variables, drop NAs for age, turn sex into a vector, normalize class

```{r}
summary(t_df)

t_clean <- t_df |> select(survived,pclass,sex,age,sib_sp,parch,fare,cabin) |> drop_na()

summary(t_clean)

age_survival <- ggplot(data=t_clean) + geom_line(aes(x=age,y=survived)) + theme_bw()


```
```{r}
surv_df <- t_df %>%
  mutate(survived = factor(survived),   ### categorical outcome variables need to be factors
         pclass   = factor(pclass)) %>% ### turn some predictors to factor
  select(-cabin, -ticket)  ### lots of NAs here - and not likely to be very helpful

```


```{r}
surv_df$gender_numeric <- ifelse(surv_df$sex == "male", 1, 0)
```

Here are some quick examples that I did.


```{r}
ggplot(surv_df, aes(x = pclass, fill = survived)) +
  geom_bar()

ggplot(surv_df,aes(x=gender_numeric,fill=survived))+geom_bar()
```



```{r}
ggplot(surv_df, aes(x = age, fill = survived)) +
  geom_histogram()

```

## Data Split

We will set aside (“partition”) a portion of the data for building and comparing our models (80%), and a portion for testing our models after we’ve selected the best one (20%). NOT the same as folds - that will happen in the training/validation step.

```{r}
### Check balance of survived column
surv_df %>%
  group_by(survived) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop = n / sum(n))
```


Tidymodels will split the data and label it for us.

```{r}
set.seed(123)

surv_split <- initial_split(surv_df, prop = 0.80,strata = survived)
  ### stratified on `survived`; training and test splits will both have ~60/40% survived = 0/1
surv_train_df <- training(surv_split)
surv_test_df <- testing(surv_split)

```

Check to make sure the data has the same proportion of splits. Why is it important to maintain the same proportion of splits?

```{r}
surv_train_df %>%
  group_by(survived) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop = n / sum(n))

surv_test_df %>%
  group_by(survived) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop = n / sum(n))
```



## Model Building

Constructing models in `tidymodels` is frighteningly simple. We tell R which kind of algorithm we want to build (model), what package the algorithm should come from (engine), and how to construct it.

```{r}
log_md <- logistic_reg() %>%
  set_engine("glm")

lasso_md<- logistic_reg(penalty = 0.037,mixture=1) %>%
  set_engine("glmnet")
```

### Data Preprocessing

We use receipes to convert our data into the format best suited to our chosen models. Basically we tell R to *consistently* transform our data between all training and testing sets. [This prevents data leakage and ensures that our models are trained on the same data](https://en.wikipedia.org/wiki/Leakage_(machine_learning)). 

We're going to build two models: a logistic regression and a lasso logistic regression model. 

```{r}
glm_rec<-recipe(survived ~ sex + pclass, data = surv_train_df)
# Could add age, fare

# steps we need to do to prepare data for lasso

lasso_rec<-recipe(survived~.,data=surv_train_df) %>%
  update_role(passenger_id, new_role = "ID") %>% #it's just a tag, it's role is just a marker
  step_rm(name,age) %>% #don't want to consider name and age since there's so much missing data
  step_unknown(all_nominal(),-all_outcomes()) |> #if you see an NA, make a new factor class called unknown
  step_dummy(all_nominal(),-all_outcomes()) |>
  step_zv(all_numeric(),-all_outcomes()) |> #if factors have no variance, get ridd of them, only mandatory elements
  step_normalize(all_numeric(),-all_outcomes()) #all numeric data should be normalized, only mandatory elements

```


### Train Model

First we create a workflow that combines all the models and the receipes to control the data. Then we use that consistent pattern to fit our model. First let's compare the models one time. Add comments to the following code chunk to describe what each step is doing. Feel free to run code.

```{r}
# First the logistic regression

log_wf <- workflow() %>%
  add_recipe(glm_rec) %>%
  add_model(log_md) 

log_fit<-log_wf %>%
  fit(surv_train_df)

log_test<-surv_test_df |>
  mutate(predict(log_fit, new_data = surv_test_df)) |> 
  mutate(predict(log_fit,new_data = surv_test_df, type='prob'))

table(log_test$survived, log_test$.pred_class)

```


Now fill in the following code chunk to fit the lasso model. Create a table (or sometimes called a confusion matrix) that shows the predicted values versus the actual values.


```{r}
lasso_wf <-workflow()|>
  add_recipe(lasso_rec)|>
  add_model(lasso_md)

lasso_fit<-lasso_wf|>
  fit(surv_train_df)

lasso_test <-surv_test_df|>
  mutate(predict(lasso_fit,new_data=surv_test_df))|>
  mutate(predict(lasso_fit,new_data=surv_test_df,type='prob'))

table(lasso_test$survived, lasso_test$.pred_class)

```

### Evaluate Performance

Measure the accuracy using the `accuracy` function from the `yardstick` package for each model.

```{r}
log_test |> 
  accuracy(truth = survived, estimate = .pred_class)

lasso_test |>
  accuracy(truth = survived, estimate = .pred_class)
```


Calculate the `ROC AUC` for each model. Use the `roc_auc` function from yardstick.

```{r}
log_test |> 
  yardstick::roc_auc(truth=survived,.pred_0) #capturing the fact that the values are different. accuracy only looks at whether or not it got the right answer, not the precision

lasso_test|>
  yardstick::roc_auc(truth=survived,.pred_0)
```


### Model Selection

One run of the model is not enough to determine which model is better. We need to run the model multiple times to determine which model is better. We can use cross-validation to determine which model is better. Instead of for loops or purrr, tidymodels as built in functions to do this for us. 



```{r}
set.seed(12)

folds<-vfold_cv(surv_train_df, v=10, strata = survived) #says how many times you want to fold data and cross-validate

log_fit_folds<- log_wf |> 
  fit_resamples(folds)

collect_metrics(log_fit_folds)

lasso_res<-lasso_wf %>%
  fit_resamples(folds)

collect_metrics(lasso_res)
```

Which model do we choose?

Let's look at the actual models to get a better understanding.

```{r}
coef<-log_fit |> 
  extract_fit_parsnip()|> 
  tidy()

lasso_fit |> #rather than me telling it to look for a specific coefficient, it takes all possible coefficients and determines which do I need and which do I not need
  extract_fit_parsnip() |> 
  tidy()
```
Lasso set many of the parameters to zero. How do you interpret this model? Why would you want to use as opposed to simple logistic regression?

### Finalize Model

We will finalize the model by fitting the model to the entire dataset. 

```{r}
final_log<-log_wf |>
  last_fit(surv_split)

useable_log<-final_log|>extract_fit_parsnip()
```


## Interpret and Visualize Results

Everything is stored in a `workflow` object. We can extract the coefficients from the logistic regression model using `extract_fit_parsnip`. The `tidy` function will make the output more readable. Describe the coefficients and what they mean. Create a clean table of the model output.

```{r}

final_log |>
  extract_fit_parsnip() |>
  tidy() |> 
  mutate(odds=exp(estimate),
         prob=odds/(1+odds))
```

Please give this a go on your own with the post-lab exercise. I will be walking around to assist.


## Tuning example

This just shows how to hypertune the `glmnet` penalty parameter. We'll cover more when we get to random forests with Yutian.

```{r}
set.seed(123)

lambda_grid <- grid_regular(penalty(), levels = 50)

lasso_md_tune<- logistic_reg(penalty = tune(),mixture=1) %>%
  set_engine("glmnet")

lasso_wf<-workflow() %>%
  add_model(lasso_md_tune) %>%
  add_recipe(lasso_rec)


set.seed(2020)
lasso_grid <- tune_grid(
  lasso_wf,
  resamples = folds,
  grid = lambda_grid
)

lasso_grid %>%
  collect_metrics()

lowest_rmse <- lasso_grid %>%
  select_best(metric="roc_auc") #use this metric
```


