---
title: "Lab 3: Post Lab"
format: html
---

## Data Load

```{r}
#import data from this website

wine_quality <- readr::read_delim("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv",delim=';')
 
```


## Design a research question

Examine the data to come up with an interesting research question based on the wine data. It is a large collection of wine attributes and quality ratings to indicate the sommlier's rating of the wine.

## Explore the data

Create some exploratory tables or charts to help you better understand possible relationships

find min and max of the score, then set that to 0 or 1
```{r}
min(wine_quality$quality)
max(wine_quality$quality)
  
```


## Use a Tidymodel framework to conduct your analysis.

```{r}
wine_df %>%
  group_by(quality) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop = n / sum(n))
```


```{r}
wine_df <- wine_quality %>%
  mutate(survived = factor(quality))|>   ### categorical outcome variables need to be factors
```


## Visualize

Present your final results and upgrade this document's YAML to a professional level quality.
